---
title: "COVID19 Delta Infections in NSW"
author: "Joe Fernando"
date: "01/09/2021"
output: 
    html_document:
        keep_md: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



### Type of Model, Assumptions and Data used in projection

__Model used__

-   The forecast model is based on the SIR Model for Spread of Disease and adopted from this blog post <https://blog.ephorie.de/covid-19-the-case-of-germany>     


__Key Assumptions__

* One the key inputs of the model is `susceptible individuals` in NSW.
* With the current vaccination rates, the model assumes at least 90% vaccination rate for 16 and older cohort in the next few weeks
* The proportion of 16 and older cohort is assumed at 81.4% and sourced from <http://www.healthstats.nsw.gov.au/Indicator/Dem_pop_age>
* At this point, no breakthrough infections are factored (which is very optimistic). 
* Moreover, the model also optimistically assumes the under 16 cohort are not at risk of infection
* We can build on the sensitivity of `susceptible individuals` later via a shiny interactive module (to follow shortly). 
* The model starts from 01Jul21      


__Data source & things to keep in mind__

-   Data for the analysis is sourced from Data.NSW and we have used the `NSW Covid19 case by location` data set. This data set differs from the numbers issued by the Premier in her daily presser due to the timing in data collection. NSW Heath reports from 8am to 8pm whilst the data set used is 00:00 Hrs to 24:00 Hrs.
-   The last data point in the data set is only fully updated the next day     



### Model output

* Lets first run the code in `code/nsw_forecast.R` to generate the projection 
```{r, echo=FALSE}

suppressMessages(source("code/nsw_forecast.R"))

```

* The code will generate two plots, one for the first 70 days to see how well actuals line-up with model and a second one for 200 days to understand full trajectory of infections. 
* There is also an additional data set of confirmed cases under the variable `nsw`. I found having this helpful as I use this scan for exposure sites close to where I live.     


#### Lets view the projection for the first 70 days:
```{r, echo=FALSE}

plot(blast_off_plot)

```

* As you can see actuals line-up nicely with the model. It makes one wonder how much of an impact targetted lockdowns have had in stopping th delta variant.
* Current Ro is at 1.14     
* Note the last data point, as mentioned above has not been updated with all of the confirmed infections, hence the low value.



#### Next lets view the potential full trajectory of the current outbreak 
```{r, echo=FALSE}

suppressMessages(source("code/nsw_forecast.R"))
plot(full_forecast_plot)

```

* Despite our optimistic assumptions, the model shows us that there is a potentially frightening few months ahead for NSW.     
* The model projects max daily infections of nearly 5,000 by 11Oct21. A frightening scenario for NSW Health!


#### Implications for NSW
* It appears NSW still has a lot of work to do to contain COVID19 transmission rates. 

* We can also use the projection to manage our own expectations of the future such as:

        + Potentially tighter restrictions on movements or a hard lockdown of some sort
        + Potentially our kids are not likely to go back to school this calendar year
        + Supply chain is likely to be interrupted, so plan ahead for provisions and Christmas gifts

* Perhaps the most important takeaway from the model is to get vaxed and to practice proper hygiene/social distancing as we are not out of the woods as yet!
        

Thanks for reading.      

Joe Fernando